Alex's Notes and Text Fragments for the JSS Manuscript

### 2021-05-06 paper/paper.Rmd
spatio-temporal -> spatiotemporal (as this post convinced me: https://english.stackexchange.com/questions/490525/spatio-or-spatial )

---
documentclass: jss
author:
  - name: Patrick Schratz
    affiliation: 'Friedrich-Schiller-University Jena'
    # affiliaton2: 'Ludwig-Maximilians-University Munich'
    # use this syntax to add text on several lines
    address: |
      | Department of Geography
      | Geographic Information Science group
    email: \email{patrick.schratz@uni-jena.de}
  - name: Marc Becker
    affiliation: 'Ludwig-Maximilians-University Munich \AND'
    address: |
      Department of Statistics
      Statistical Learning and Data Science group
    # To add another line, use \AND at the end of the previous one as above
  - name: Michel Lang
    affiliation: TU Dortmund University
    address: |
      | Faculty of Statistics
  - name: Alexander Brenning
    affiliation: 'Friedrich Schiller University Jena'
    address: |
      | Department of Geography
      | Geographic Information Science group
    email: \email{alexander.brenning@uni-jena.de}
title:
  formatted: "\\pkg{mlr3spatiotempcv}: Spatiotemporal resampling methods for \\pkg{mlr3}"
  # If you use tex in the formatted title, also supply version without
  plain:     "mlr3spatiotempcv: spatiotemporal resampling methods for mlr3"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The R package [mlr3spatiotempcv](https://mlr3spatiotempcv.mlr-org.com) is an extension package of the machine learning framework [mlr3](https://mlr3.mlr-org.com) [@mlr3].
  It adds support for spatiotemporal resampling methods and respective visualization methods.
  Spatiotemporal resampling methods are needed to account for the existence of different autocorrelation types (spatial autocorrelation (SAC), temporal autocorrelation (TAC) or spatio-temporal autocorrelation (STAC)) in datasets.
  These autocorrelation types cause biased performance estimates of varying magnitude in cross-validation when ignored.
  At the time of writing, various R packages implementing different spatiotemporal partitioning strategies exist: CAST [@cast], blockCV [@blockCV] or sperrorest [@sperrorest].
  There are some resampling methods used in scientific studies for which the source code is either not available or a different programming language has been used [@roberts2017; @pohjankukka2017]

@@

Spatial and spatio-temporal prediction tasks are common in applications ranging from environmental sciences to archaeology and epidemiology. 
While sophisticated mathematical frameworks have long been developed in spatial statistics to characterize predictive uncertainties under well-defined mathematical assumptions such as intrinsic stationarity [e.g., cressie2015], computational estimation procedures have only been proposed more recently to assess predictive performances of spatial and spatio-temporal prediction models [@brenning2005; @sperrorest; @pohjankukka2017; @roberts2017].

@@

While it is not the purpose of this paper to recommend specific resampling schemes for specific use cases, this example may suffice to motivate the use of spatial and spatio-temporal cross-validation techniques, and the need for a unified framework and computational toolbox that accommodate a variety of prediction tasks that may be applicable to a broad range of prediction tasks.
\pkg{mlr3spatiotempcv} is such a toolbox.

This toolbox, implemented as an open-source R package, builds upon and generalizes several existing toolboxes that have been developed in recent years for more specific settings (see Table \@ref(tab:sptcv-methods)). 
The earliest of these implementations is the \pkg{sperrorest} package [@sperrorest], which has, however, not been integrated into established machine-learning frameworks such as \pkg{mlr} or \pkg{caret}, and lacks support of temporal prediction tasks. 
Other more recent implementations include \pkg{blockCV} for (what for?) [ref @blockCV; @rest2014] and \pkg{CAST} for spatio-temporal prediction tasks in a \pkg{caret} framework [@cast; @meyer2018a].

Thus, \pkg{mlr3spatiotempcv} implements for the first time a comprehensive state-of-the-art compilation of spatial and spatio-temporal partitioning schemes that is well-integrated into a major machine-learning framework in R, the \pkg{mlr3} ecosystem [@mlr3]. 
This package is furthermore equipped with a variety of two- and three-dimensional diagram types. Our hope is that implementation will facilitate reproducible geospatial modelling and code-sharing across a broad range of application domains.

# Spatial and Spatio-temporal Cross-validation

@@

With the increased awareness of the importance of spatial and spatio-temporal resampling strategies and the growing popularity of R in environmental modelling and geocomputation, it is important to equip machine-learning frameworks such as \pkg{mlr} with suitable algorithms. 
In this context, \pkg{mlr3} stands out as a unified, object-oriented and extendable framework designed to accommodate numerous machine-learning tasks with a variety of learners, feature and model selection tools, and model assessment capabilities [@mlr3; mlr3book]. 
All of these are supported by advanced visualization tools, which are particularly important in a spatial and spatio-temporal setting.

@@

## Comparing spatio-temporal resampling methods

@@

<!-- missing: Guidelines how to choose a resampling method -->
While various spatio-temporal resampling methods exists nowadays to account for spatio-temporal autocorrelation in modeling, there is still lots of uncertainties which method should be used for which dataset.
The answer to this questions is not trivial and needs more research.
One option could be to propose a checklist of dataset characteristics which would make a dataset more suitable for specific methods that could help users making an appropriate and objective decision.

@@

We believe that making resampling methods available to widely used modeling frameworks is more benefitial than implementing such methods standalone in an R package.
The latter often require special structured objects as inputs and cannot directly be re-used in susequent modeling steps of user workflows.
This is not an issue specific to spatio-temporal resampling methods but a general philosophy of the authors.

Spatio-temporal cross-validation is still not yet a paradigm that is completely established in scientfic workflows even though it is discussed intensively for more than a decade now.
Making most available methods easily accessible to users is an important step to foster the acceptance of such in the community and lead to less biased cross-validation results in environmental and ecological studies.

>>>>

---
documentclass: jss
author:
  - name: Patrick Schratz
    affiliation: 'Friedrich-Schiller-University Jena'
    # affiliaton2: 'Ludwig-Maximilians-University Munich'
    # use this syntax to add text on several lines
    address: |
      | Department of Geography
      | Geographic Information Science group
    email: \email{patrick.schratz@uni-jena.de}
  - name: Marc Becker
    affiliation: 'Ludwig-Maximilians-University Munich \AND'
    address: |
      Department of Statistics
      Statistical Learning and Data Science group
    # To add another line, use \AND at the end of the previous one as above
  - name: Michel Lang
    affiliation: TU Dortmund University
    address: |
      | Faculty of Statistics
  - name: Alexander Brenning
    affiliation: 'Friedrich Schiller University Jena'
    address: |
      | Department of Geography
      | Geographic Information Science group
    email: \email{alexander.brenning@uni-jena.de}
title:
  formatted: "\\pkg{mlr3spatiotempcv}: Spatiotemporal resampling methods for \\pkg{mlr3}"
  # If you use tex in the formatted title, also supply version without
  plain:     "mlr3spatiotempcv: spatiotemporal resampling methods for mlr3"
  # For running headers, if needed
  # short:     "\\pkg{foo}: A Capitalized Title"
abstract: >
  The R package [mlr3spatiotempcv](https://mlr3spatiotempcv.mlr-org.com) is an extension package of the machine learning framework [mlr3](https://mlr3.mlr-org.com) [@mlr3].
  It adds support for spatiotemporal resampling methods and respective visualization methods.
  Spatiotemporal resampling methods are needed to account for the existence of different autocorrelation types (spatial autocorrelation (SAC), temporal autocorrelation (TAC) or spatiotemporal autocorrelation (STAC)) in datasets.
  These autocorrelation types cause biased performance estimates of varying magnitude in cross-validation when ignored.
  At the time of writing, various R packages implementing different spatiotemporal partitioning strategies exist: CAST [@cast], blockCV [@blockCV] or sperrorest [@sperrorest].
  There are some resampling methods used in scientific studies for which the source code is either not available or a different programming language has been used [@roberts2017; @pohjankukka2017]

@@

Spatial and spatiotemporal prediction tasks are common in applications ranging from environmental sciences to archaeology and epidemiology. 
While sophisticated mathematical frameworks have long been developed in spatial statistics to characterize predictive uncertainties under well-defined mathematical assumptions such as intrinsic stationarity [e.g., cressie2015], computational estimation procedures have only been proposed more recently to assess predictive performances of spatial and spatiotemporal prediction models [@brenning2005; @sperrorest; @pohjankukka2017; @roberts2017].

@@

While it is not the purpose of this paper to recommend specific resampling schemes for specific use cases, this example may suffice to motivate the use of spatial and spatiotemporal cross-validation techniques, and the need for a unified framework and computational toolbox that accommodate a variety of prediction tasks that may be applicable to a broad range of prediction tasks.
\pkg{mlr3spatiotempcv} is such a toolbox.

This toolbox, implemented as an open-source R package, builds upon and generalizes several existing toolboxes that have been developed in recent years for more specific settings (see Table \@ref(tab:sptcv-methods)). 
The earliest of these implementations is the \pkg{sperrorest} package [@sperrorest], which has, however, not been integrated into established machine-learning frameworks such as \pkg{mlr} or \pkg{caret}, and lacks support of temporal prediction tasks. 
Other more recent implementations include \pkg{blockCV} for (what for?) [ref @blockCV; @rest2014] and \pkg{CAST} for spatiotemporal prediction tasks in a \pkg{caret} framework [@cast; @meyer2018a].

Thus, \pkg{mlr3spatiotempcv} implements for the first time a comprehensive state-of-the-art compilation of spatial and spatiotemporal partitioning schemes that is well-integrated into a major machine-learning framework in R, the \pkg{mlr3} ecosystem [@mlr3]. 
This package is furthermore equipped with a variety of two- and three-dimensional diagram types. Our hope is that implementation will facilitate reproducible geospatial modelling and code-sharing across a broad range of application domains.

# Spatial and Spatiotemporal Cross-validation

@@

With the increased awareness of the importance of spatial and spatiotemporal resampling strategies and the growing popularity of R in environmental modelling and geocomputation, it is important to equip machine-learning frameworks such as \pkg{mlr} with suitable algorithms. 
In this context, \pkg{mlr3} stands out as a unified, object-oriented and extendable framework designed to accommodate numerous machine-learning tasks with a variety of learners, feature and model selection tools, and model assessment capabilities [@mlr3; mlr3book]. 
All of these are supported by advanced visualization tools, which are particularly important in a spatial and spatiotemporal setting.

@@

## Comparing spatiotemporal resampling methods

@@

<!-- missing: Guidelines how to choose a resampling method -->
While various spatiotemporal resampling methods exists nowadays to account for spatiotemporal autocorrelation in modeling, there is still lots of uncertainties which method should be used for which dataset.
The answer to this questions is not trivial and needs more research.
One option could be to propose a checklist of dataset characteristics which would make a dataset more suitable for specific methods that could help users making an appropriate and objective decision.

@@

We believe that making resampling methods available to widely used modeling frameworks is more benefitial than implementing such methods standalone in an R package.
The latter often require special structured objects as inputs and cannot directly be re-used in susequent modeling steps of user workflows.
This is not an issue specific to spatiotemporal resampling methods but a general philosophy of the authors.

Spatiotemporal cross-validation is still not yet a paradigm that is completely established in scientfic workflows even though it is discussed intensively for more than a decade now.
Making most available methods easily accessible to users is an important step to foster the acceptance of such in the community and lead to less biased cross-validation results in environmental and ecological studies.
